---
title: "Untitled"
format: html
---


# Exercise 

1. Create a create_barchart() function that plots a bar chart of a given variable.

Bonus: make it so that bars are sorted by size (tip: use a function from the fct_ family)
```{r}
create_barplot <- function(data, variable_to_plot){
  data %>% 
    mutate({{variable_to_plot}} := fct_rev(fct_infreq({{variable_to_plot}}))) %>% 
    ggplot(aes(y = {{variable_to_plot}})) +
    geom_bar() +
    theme_light()
}

create_barplot(iris, Species)
```


2. Change the create_scatterplot() function so that it has a title that describes which variables are plotted.

Tip: this requires treating the name of the objects as a string. Perhaps a quick google search can help you.

```{r}
library(tidyverse)
library(glue)

create_scatter_plot <- function(data, variable1, variable2) {
  variable_name1 <- ensym(variable1)
  variable_name2 <- ensym(variable2)
  
  data %>% 
    ggplot(aes({{variable1}}, {{variable2}})) +
    geom_point() +
    theme_light() +
    labs(title = glue("Scatterplot between {variable_name1} and {variable_name2}"))
}

```

```{r}
create_scatter_plot(mtcars, mpg, cyl)
```

