---
title: "R Markdown Example"
author: "Your Name"
date: "" 
output:
  html_document:
    toc: true
    toc_float:
      collapsed: true
    toc_depth: 4
    theme: flatly
    code_folding: show
    number_sections: false
    df_print: paged
  editor_options:
    chunk_output_type: console
---

# Introduction

This is an example of an Rmarkdown document. Rmarkdown is a powerful tool for combining code, text, and visualizations in a single document. It allows you to create documents that are reproducible, dynamic, and interactive.

It is super versatile, and a lot of formatting and styling options can be specified using the YAML header (the section at the very top of this document).

# Code Chunks

R Markdown documents are made up of code chunks and markdown text. Code chunks are blocks of R code that can be executed and their output can be included in the document. Here is an example of a code chunk:

```{r}
x <- rnorm(100)
mean(x)
```

The `{r}` at the beginning of the code chunk indicates that this is an R code chunk. When you run the code, the output will be included in the document. In this case, the output would be the mean of 100 random normal variables.

You can run all code within a chunk by clicking on the small green arrow pointing to the right at the top of a code chunk ("run current chunk").

You can also run all code above a chunk by clicking on the downward arrow ("Run all chunks above"). This is excellent for resuming work when reopening a project, and for checking for reproducibility of your results (make sure to restart R before doing so).

# Markdown Text

In addition to code chunks, you can also include plain text in an R Markdown document using markdown syntax. Markdown is a simple formatting syntax that allows you to add emphasis, headings, lists, and other formatting to your text.

For example, you can create a heading using a `#` symbol:

\# Heading 1\
\## Heading 2\
\### Heading 3

In the knitted document, this converts to:

# Heading 1

## Heading 2

### Heading 3

You can also create a bullet list using `-`:

-   item 1
-   item 2
-   item 3

# Visualizations

R Markdown allows you to include plots and visualizations in your document. Here is an example of a plot generated using R:

```{r, echo = FALSE}
x <- rnorm(100)
y <- rnorm(100)

plot(x, y)
```

# Working with Rmarkdown

R markdown is great for working with data related projects, as it allows you to work efficiently and keep your code organized almost automatically. Plus, you get a report of your analysis (the "knitted" version of your markdown file) for free.

Here's an example of how to work with the EU data

# Setup

First, we specify a few default `knitr` options, using the `knitr::opts_chunk$set()` function. "knitr" is the package that deals with the knitting of rmarkdown documents to the specified format (e.g., html, pdf, word, ...). You will not need to deal with it much, but understanding a few concepts will help you produce prettier documents, and allow you to make better use of its customization potentials.

```{r setup, include=TRUE}
# the following code will set default settings for how code chunks should be knitted
knitr::opts_chunk$set(
  echo = TRUE, # code chunks will appear in the knitted document
  message = FALSE, # messages from code you run will not appear in the knitted document
  warning = FALSE, # warnings from code you run will not appear in the knitted document
  fig.width = 12, # figures will have a default width of 12 inches
  fig.height = 7, # figures will have a default height of 7 inches
  cache = TRUE # code chunks will be cached when they are run. This allows speedier re-knitting of documents
)


```

Then, we load the necessary packages. It is good practice to load all packages you need for your analysis / project at the start of the script, so others can see whether they need to install something before they can run the full script.

```{r}
library(tidyverse)
library(skimr)
```

# Loading and Preprocessing Data

## Loading Data

Here, we load the raw data from URL.

```{r}
raw_data <- read_csv("https://comparativeagendas.s3.amazonaws.com/datasetfiles/ECC_7.0_1975-2014_1.csv")
```

## Preprocessing

Here, we do some preprocessing of the data. For instance, we might get rid of a bunch of colums in the raw dataset that we do not need

```{r}
cleaned_dat <- raw_data |>
  select(-starts_with("filter"))
```

## Data overview

Here, we give an overview of the data in the cleaned dataset. We're using the `skimr` package, which does this automatically for you.

```{r}
skimr::skim(cleaned_dat)
```

# Analysis

Here, we could start with our analysis.

For instance, we could show some kind of plot. The following plot shows the count of major topics in the EU_CAP dataset by year.

```{r}
growth_data <- cleaned_dat |>
  # group dataset by year 
  group_by(year) |>
  
  # count the number a major topic occurs per year
  count(majortopic) |>
  
  # make "majortopic" a factor variable to help with plotting
  mutate(majortopic = as.factor(majortopic))


growth_data |>  
  # plot the data using ggplot
  ggplot(aes(year, n, colour = majortopic)) +
  geom_line() +
  theme_light()
```

Then, we could test some kind of model. Maybe how much a given topic has grown over the years (this is of course not how you would model this, but just for illustration):

```{r}
topic_growth_model <- lm(n ~ year + majortopic,
                         data = growth_data)

summary(topic_growth_model)
```

I recommend the `sjPlot::tab_model()` function for outputting nicely formatted regression tables:

```{r}
sjPlot::tab_model(topic_growth_model)
```

# Find out more

## External resources

There's a ton of stuff to learn with rmarkdown. Get inspiration here: <https://rmarkdown.rstudio.com/index.html>

## Practical tips

### YAML header options

Here's the YAML header options I like to specify for my projects (see also above):

    title: "Heading of your document"
    author: "Your Name"
    date: "Date" 
    output:
      html_document:
        toc: true
        toc_float:
          collapsed: true
        toc_depth: 4
        theme: flatly
        code_folding: hide
        number_sections: true
        df_print: paged
      editor_options:
        chunk_output_type: console

### HTML Styling

In addition, I like to add in this special html block for styling (view rmd document to see this chunk):

```{=html}
<style type="text/css">
.main-container {
  max-width: 1400px !important;
  width: 1400px !important;
  margin-left: auto;
  margin-right: auto;
}
body {
  font-size: 14pt;
  max-width: 1400px !important;
  width: 1400px;
}
p {
  padding-bottom: 12px;
  padding-top: 12px;
  line-height: 1.15em;
}
</style>
```
### Tabulated content {.tabset}

Working with tabulated content. You can do so by adding `{.tabset}` to your heading.

Every subheading on that level will now be in a tab, like so:

#### Tab 1

Some content for tab 1

#### Tab 2

Different content for tab 2

#### Tab 3

Yet other content content for tab 3

###  {.unnumbered}

Neat, right?
