---
title: "The Tidyverse - A Recap"
format: 
  revealjs:
    theme: [simple, ../assets/custom.scss]
    df-print: paged
    code-line-numbers: true
editor: source
execute: 
  echo: true
link-external-newwindow: true
---

```{r}
#| echo: false

library(tidyverse)
```

# Day 2

## Questions before we begin?

## Data cleaning & manipulation in the [tidyverse]{.blue_highlight}

1.  **Basics** - mutate, summarise, group_by, pivot_longer, pivot_wider, ...
2.  **Advanced concepts** - across, rowwise, tidyselect methods, ...
3.  **Text data ** - Regular expressions, {Stringr}, {Glue}
4.  **Practice** :)

# Recap of most important tidyverse syntax and functions

## Bread and butter functions

-   `mutate()`
-   `group_by()`
-   `summarise()`
-   `filter()`
-   `select()`

## New as of dplyr 1.1.2 {.smaller}

Since dplyr 1.1.2, you can use the "[**.by**]{.blue_highlight}" argument to use `group_by()` within selected dplyr verbs

. . .

```{r}
#| output-location: fragment
mtcars %>% 
  mutate(avg_mpg = mean(mpg), .by = cyl) %>% 
  head()
```

. . .


Also works with `summarise(.by = )`, `reframe(.by = )`, `filter(.by = )`, `slice(.by = )`

## Reshaping

-   `pivot_wider()` (formerly: `spread()`)
-   `pivot_longer()` (formerly: `gather()`)

## Joining

-   `left_join()`
-   `right_join()`
-   `inner_join()`
-   `full_join()`

## The "Pipe"

-   `%>%()`
-   `|>()`

. . .

```{r}
#| eval: false
The_pipe %>% 
  allows_us() %>% 
  to_string_together() %>% 
  multiple_functions()

# this is the same as
multiple_functions(to_string_together(allows_us(The_pipe)))
```


# Practice

## Let's play {.scrollable .smaller}

Explore the chocolate dataset from tidytuesday, see [here for documentation](https://github.com/rfordatascience/tidytuesday/blob/master/data/2022/2022-01-18/readme.md)


```{r}
chocolate <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-01-18/chocolate.csv')
```

- Use `group_by` and `summarise` functions to explore mean ratings for different countries, beans, years...
- use the pipe!
- Plot the data in different ways


. . .

#### Bonus:

**Do some data cleaning:** the ingredients and characteristics column encode several observations. Separate out these columns, and make sure data is in long format :)


